Interpretable Neural Networks using EAGGA

### 2025-03-13 paper/main.tex
finish MOO bullets

\subsection{Multi-Objective Optimization}
- in most practical use cases one doesn't just want to optimise for performance but e.g. also for interpretability using some proxy metrics
- formal definition
\subsubsection{Pareto-optimality}
\subsubsection{A-priori}
- requires specifying trade-off between objectives a-priori
- simplest approach: different versions of \textbf{scalarization}
  * weighted sum of objective functions
    - $\min_{x\in\mathcal{X}} w_i f_i(x)$ with $\sum_{i=1}^k w_i=1$ and $w_i>0,\forall i=1,...,k$
    - \citep[p. 11]{10.1145/3610536}
  * Chebyshev: provably converges to Pareto front
\subsubsection{A-posteriori}
- problem a priori
  * either restrict search space to ``enforce our will'' (e.g. only use 50\% of features), optimise only prediction performance + use this
  * or leave search space unrestricted but adjust loss function to incorporate multiple objectives + take optimum from there
  * in either case: no knowledge of interplay between HP config + performance on all objectives
  * a-posteriori evaluates configs just as a-priori, but keeps track of multiple ``best'' (non-dominated) solutions -> makes relationship between HP config + objectives visible
- NSGA-II (MOEA)
  * non-dominated sorting
  * crowding distance
  -> give intuition why non-dom-sort (asc) + crowding dist (desc) makes sense for selecting individuals

>>>>

\subsection{Multi-Objective Optimization}
- in most practical use cases one doesn't just want to optimise for performance but e.g. also for interpretability using some proxy metrics
- formal definition
  * vector of objectives $c_1,c_2,...,c_m=\boldsymbol{c}:\Lambda\rightarrow\mathbb{R}^m$
  * goal: minimise vector
  * \citep[p. 11]{10.1145/3610536}
\subsubsection{Pareto-optimality}
- problem: usually conflicting objectives, i.e. minimisation of $\boldsymbol{c}$ not possible along all dimensions
- thus aim to find trade-off solutions of non-dominated points
- we say a point dominates another
  * if there is no other point strictly better in at least one dimension and better or equal in the remaining ones
  * formally: $\boldsymbol\lambda$ dominates $\boldsymbol\lambda'$ ($\boldsymbol\lambda\prec\boldsymbol\lambda'$) if and only if
    $\forall i\in\{1,...,m\}:c_i(\boldsymbol\lambda) \le c_i(\boldsymbol\lambda') \wedge \exists j\in\{1,...,m\}:c_j(\boldsymbol\lambda) < c_j(\boldsymbol\lambda')$
  * \citep[pp. 7f]{10.1145/3610536} + \citep[pp. 198f]{genetic_algos}
- Pareto set: set of nondominated points $\mathcal{P}:=\{\boldsymbol\lambda\in\Lambda|\nexists\boldsymbol\lambda'\in\Lambda\text{ s.t. }\boldsymbol\lambda'\prec\boldsymbol\lambda\}$
- Pareto front: image of nondominated points
- goal: find set of nondominated points $\hat{\mathcal{P}}$ approximating true Pareto set $\mathcal{P}$ well
\subsubsection{A-priori}
- requires specifying trade-off between objectives a-priori, will outline 2 popular approaches
- e.g. different versions of \textbf{scalarization}
  * weighted sum of objective functions
    - $\argmin_{\boldsymbol\lambda\in\Lambda} w_i c_i(\boldsymbol\lambda)$ with $\sum_{i=1}^k w_i=1$ and $w_i>0,\forall i=1,...,k$
    - drawbacks
      * solution sensitive to weights
      * different users might have different opinions on weights \citep[chap. 3.1]{NSGA}
    - \citep[p. 11]{10.1145/3610536}
  * $\epsilon$-constraint
    - translate all but one objective into constraint, then optimise remaining objective under subject to the constraints
    - $\text{w.l.o.g.} \argmin_{\boldsymbol\lambda\in\Lambda} c_1(\boldsymbol\lambda) \text{ s.t. } c_2(\boldsymbol\lambda)\le\epsilon_2,...,c_m(\boldsymbol\lambda)\le\epsilon_m$
    - conceputally similar to weighted sum: also sensitive to constraints, must be chosen sensibly
    - \citep[p. 12]{10.1145/3610536}
- \textbf{lexicographic method}
  * define priorisation of objectives
  * greedily optimise objectives in order of priority, constraint to the solutions of the already optimised (higher-priority) objectives
  * again very dependent on user-defined priorisation
  * \citep[p. 13749]{lexicographic_MOO}  % TODO: perhaps find a different source more related to ML
\subsubsection{A-posteriori}
- problem a priori
  * either restrict search space to ``enforce our will'' (e.g. only use 50\% of features), optimise only prediction performance + use this
  * or leave search space unrestricted but adjust loss function to incorporate multiple objectives + take optimum from there
  * in either case: no knowledge of interplay between HP config + performance on all objectives
  * in practical applications, it is oftentimes useful to make the decision of which point of the Pareto set to use a-posteriori, e.g. if only a slight decrease in
    one objective translates to a significant improvement in another that would have been missed if the problem was optimised using a-priori methods
  * a-posteriori evaluates configs just as a-priori, but keeps track of multiple ``best'' (non-dominated) solutions -> makes relationship between HP config + objectives visible
- aside from usualy baselines grid and random seach there are multi-objective BO adaptations, mainly using either of two approaches
  (1) fit single surrogate model on scalarised objectives, e.g. ParEGO, which employs the augemented Tchebycheff function as scalarisation to ensure the Pareto front is
      explored sufficiently \citep[pp. 15f]{10.1145/3610536} + \citep[pp. 54-56]{ParEGO}
  (2) fit one surrogate per objective, then
    - either one acquisition function per objective -> return set of promising next candidates
    - or one overall acquisition function aggregating surrogates, e.g. EHI, maximises expected improvement of hypervolume \citep[p. 16]{10.1145/3610536} + \citep[pp. 8f]{EHI}
- lastly, there is also multi-objective EA (MOEA) algorithms to explore Pareto front, one of which is NSGA-II (Nondominated Sorting Genetic Algorithm)
  * improves upon predecessor NSGA \citep{NSGA} by making it parameterless, ensuring elitism, and reducing computational complexity of ranking of individuals \citep[p. 182]{NSGA_II}
  * uses all the regular operators, i.e. reproduction, mutation, crossover can be used from single-objective EA (SOEA)
  * difference to SOEA: ranking of individuals (SOEA uses scalar fitness, MOEA mutliple objectives)
    - ranking mechanism based on 2 parts
      * non-dominated sorting, ensures elitism
        - iterative procedure to rank individuals in population by their fronts
        (1) determine pareto front, assign rank 0
        (2) remove pareto front from population
        (3) if individuals left, repeat from 1, increment rank \#
        - \citep[p. 201]{genetic_algos} + \citep[pp. 183f]{NSGA_II}
      * crowding distance, ensures sufficient diversity, i.e. exploration of pareto front
        - assigns score depending on how crowded area around individual in objective space is
        - crowding distance of an individual = mean side length of cuboid spanned by its nearest neighbours as vertices in objective space
        - individuals without two neighbour in any dimension are assigned infinitely high distance value
        - \citep[p. 185]{NSGA_II}
        -> the less the crowding distance, the more ``crowded'' an area is by other individuals
      -> rank individuals by nds front rank (ascending) + crowding distance (descending) as tie breaker
  * ranking used for selecting $\mu$ best individuals to keep for next generation + for reproduction: via binary tournament solution: select two random individuals,
    pick best w.r.t. nds + cd for offspring creation (mutation / crossover)
  * using nds for ranking / fitness makes intuitive sense, but why cd?
    - goal of MOO: not simply optimise HV, but approx true Pareto front well
    - only having individuals representing one area in objective space makes Pareto front estimate very ``unstable'': removing this area (i.e. solutions form that area)
      would ``collapse'' entire front
    - instead having multiple areas be represented makes the estimate ``stable'': removing one area wouldn't impact Pareto front estimate a lot
    % TODO: reference?
